### 2019-01-18 referee_response.txt
Final version of the response to the referee.

Dear PRD Editor,

Thank you for sending us the report of the referee, whom we thank for
carefully reading our manuscript and recommending publication. Following 
the recommendation of the referee, we have made a few changes in the 
abstract and in the conclusions to address his/her request to point 
out the effects attributable to the scalar field. The changes made in 
the manuscript are highlighted in boldface, to facilitate identification.

       Best wishes,

       The authors

> ----------------------------------------------------------------------
> Report of the Referee -- DY12222/Gimeno-Soler
> ----------------------------------------------------------------------

> The paper describes an interesting and potentially important problem
> concerning the structure of accretion disks around exotic black holes
> modified by the presence of a gravitationally strong scalar field. The
> authors focus on the semi-analytic models of magneto-static tori,
> which under some conditions may capture properties of such disks, and
> build a grid of models exploring the parameter space. This is a
> meticulous work of good quality, which I am happy to recommend for
> publication in Physics.Rev.D.

> There is however one issue I would like the authors to address. The
> first question one would expect to be answered in such a paper is how
> strong, robust and unmistakable the effects attributable to the scalar
> field are. Is there a smoking gun, so to speak? Unfortunately the
> authors avoid addressing this question both in the Abstract and in the
> Conclusions, which is rather puzzling.

We agree with the referee and therefore we have added the following 
text to the abstract and conclusions:

Abstract:

"The sequences of magnetized, equilibrium disks around Kerr black holes 
with scalar hair discussed in this study are morphologically and 
thermodynamically different than their Kerr black hole counterparts, 
namely their vertical size is larger, the high-density central region 
is more extended, and the fluid is more relativistic. Therefore, we 
expect significant differences to appear when these sequences are 
used as initial data for numerical relativity codes to investigate 
their dynamical (non-linear) stability and used in tandem with 
ray-tracing codes to obtain synthetic images of black holes (i.e. 
shadows) in astrophysically relevant situations where the light 
source is provided by an emitting accretion disk."

Conclusions:

"In particular, the morphological and thermodynamical differences we 
observe when comparing the solutions for KBHsSH with those for disks 
around Kerr BHs (i.e. a larger vertical size of the disks, a more 
extended high-density central region, and a more relativistic fluid) 
are expected to yield significantly different results when their 
shadows are computed. Moreover, the stability properties of such disks 
under perturbations are expected to be affected by the presence of the 
gravitational well produced by the scalar hair."